---
title: "Digital Equity Information - Dane County Wisconsin"
author: "Deanna Schneider"
date: "2023-05-04"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, error = F, message = F)
```

The following information is based on the most recent 5-year American community survey data (ending with 2021).


```{r results='hide'}
library(tidycensus)
library(tidyverse)
library(tigris)
library(leaflet)
library(tmap)
tmap_mode('view')#switch to view for interactive maps, plot for static
options(tigris_use_cache = TRUE)

rappdirs::user_cache_dir("tigris")
vars = load_variables(year = 2021, dataset='acs1')

vars %>% filter(str_detect(name, 'B28003'))
```
```{r}
#number of use residents born in Mexico, by state from the 2020 ACS
computer_info_tract <- get_acs(
  geography = "block group", 
  variables = c("B28003_002", "B28003_003", 'B28003_004', 'B28003_005', 'B28003_006'),
  summary_var = 'B28003_001',
  #year = 2021,
  state=55,
  county=025, 
  geometry = TRUE
) %>%
  mutate(percent = 100 * (estimate/summary_est))

computer_info_dane <- get_acs(
  geography = "county", 
  variables = c("B28003_002", "B28003_003", 'B28003_004', 'B28003_005', 'B28003_006'),
  summary_var = 'B28003_001',
  #year = 2021,
  state=55,
  county=025, 
  geometry = TRUE
) %>%
  mutate(percent = 100 * (estimate/summary_est))

computer_info_wi <- get_acs(
  geography = "state", 
  variables = c("B28003_002", "B28003_003", 'B28003_004', 'B28003_005', 'B28003_006'),
  summary_var = 'B28003_001',
  #year = 2021,
  state='WI',
  geometry = TRUE
) %>%
  mutate(percent = round(100 * (estimate/summary_est), 2))


get_info_graphic <- function(var, emoji, title){
  dane_percent_of_computers <- paste0(round(computer_info_dane %>% filter(variable == var) %>% pull(percent), 1), '%')
wi_percent_of_computers <- paste0(round(computer_info_wi %>% filter(variable == var) %>% pull(percent), 1), '%')

# create two simple tiles
key_tiles <- tibble(
  x = c(2, 8.5),
  y = c(6.5, 6.5),
  h = c(4.25, 4.25),
  w = c(6.25, 6.25),
  val = c(dane_percent_of_computers, wi_percent_of_computers),
  text = c("Dane County", "Wisconsin"),
  icon = c(emojifont::emoji(emoji),
            emojifont::emoji(emoji)),
  color = factor(c(1, 2))
) 

# plot two key highlight tiles
ggplot(data = key_tiles, 
       aes(x, y, height = h, width = w, label = text)) +
  geom_tile(aes(fill = color)) +
  geom_text(color = "white", fontface = "bold", size = 16, family = "Avenir Next",
            aes(label = val, x = x - 2.9, y = y + 1.5), hjust = 0) +
  geom_text(color = "white", fontface = "bold", size = 6, family = "Avenir Next",
            aes(label = text, x = x - 2.9, y = y - 1.5), hjust = 0) +
  coord_fixed() +
  scale_fill_manual(values = c("#8cc152", "#226d93")) +
  geom_text(size = 30, 
            aes(label = icon, 
                family = "Arial", 
                x = x + 1.5, y = y + 0.3), 
            alpha = 0.25) +
  theme(plot.margin = unit(c(-0.30,0,0,0), "null")) + # remove margin around plot
  theme_void() +
  labs(caption = "Source: American Community Survey 5-year Estimates", title=title) +
  guides(fill = FALSE)
  
}

get_block_group_plot <- function(var, title, incHist=T) {
  data <- computer_info_tract %>% filter(variable == var)

  dane_pal <- c('#d8e2ea', '#a9bfd1', '#7ca0b9', '#5385a6', '#226d93')
  dane_pal_green <- c('#f7fbf3', '#e6f1d7', '#cfe4b3', '#b0d486', '#8cc152')
  
  tm_shape(data) + tm_polygons(col='percent',
                                         style='jenks', 
                                         n=5,
                                         palette=dane_pal_green,
                                         title='ACS Estimates', 
                                         legend.hist = incHist,
                               alpha = 0.75
                                         ) + 
      tm_layout(main.title = title,
              frame = FALSE,
              legend.outside = TRUE,
              legend.hist.width = 5
              ) +
    tm_basemap(leaflet::providers$OpenStreetMap)

  
}
```

## Households with a Computer
Most households in Dane County have a computer, and the percent of households in Dane County with a computer exceeds the percent of households with a computer for the state of Wisconsin. But, block group data reveals a disparity in percent of households with a computer by neighborhood, with certain neighborhoods within the county have a much lower percentage of the household owning a computer.

```{r results='hide'}
emojifont::search_emoji("phone")
```


```{r}
get_info_graphic('B28003_002', 'desktop_computer', "Households with a Computer")
get_block_group_plot('B28003_002', "Percent Households with a Computer")

```

## No Internet Access
When viewed at the county or state level, most households that have a computer have some form of internet access. But pockets exist within Dane County in which a much higher percent of households do not have internet access, even while having a computer.

```{r}
get_info_graphic('B28003_005', 'mobile_phone_off', "Households with No Internet Access")
get_block_group_plot('B28003_005', "Percent Households with Computer and No Internet", F)
```



## Dial-up Access Only
In both Dane County and Wisconsin, the percentage of households that have a computer, but only have dial-up access, is extremely low. A greater disparity is once again seen at the block group level, with some rural and urban block groups showing higher percentages of dial-up only access.

```{r}
get_info_graphic('B28003_003', 'telephone', "Households with Dial-up Access Only")
get_block_group_plot('B28003_003', "Percent Households with Computer and Dial-up Access Only", F)
```


## Broadband Access
Dane County exceeds Wisconsin in the percentage of households that have a computer and broadband access. But again, the distribution of that broadband access throughout the county is inconsistent.

```{r}
get_info_graphic('B28003_004', 'computer', "Households with Broadband Access")
get_block_group_plot('B28003_004', "Percent Households with Computer and Broadband Access", F)

```

